\chapter*{Statutory Declaration in Lieu of an Oath}
\thispagestyle{empty}
\begin{small}
\noindent The present translation is for your convenience only.\\
\noindent Only the German version is legally binding.
\vspace{2\baselineskip}

\noindent I hereby declare in lieu of an oath that I have completed the present \THESISTYPE's entitled
\begin{center}
  \TITLEENG
\end{center}
\noindent independently and without illegitimate assistance from third parties. I have use no other than the specified sources and aids. In case that the thesis is additionally submitted in an electronic format, I declare that the written and electronic versions are fully identical. The thesis has not been submitted to any examination body in this, or similar, form.

\vfill
\noindent \textbf{Official Notification}
\vspace{1\baselineskip}

\noindent \textbf{Para. 156 StGB (German Criminal Code): False Statutory Declarations}

\noindent Whosoever before a public authority competent to administer statutory declarations falsely makes such a declaration or falsely testifies while referring to such a declaration shall be liable to imprisonment not exceeding three years or a fine.
\vspace{1\baselineskip}

\noindent \textbf{Para. 161 StGB (German Criminal Code): False Statutory Declarations Due to Negligence}

\noindent (1) If a person commits one of the offences listed in sections 154 to 156 negligently the penalty shall be imprisonment not exceeding one year or a fine.

\noindent (2) The offender shall be exempt from liability if he or she corrects their false testimony in time. The provisions of section 158 (2) and (3) shall apply accordingly.
\vspace{1\baselineskip}

\noindent I have read and understood the above official notification.
\end{small} 